Hydras \& Co.

### 2023-06-15 doc/chapter-fol.tex
small changes in documentation

This chapter is devoted to the presentation to  data structures for representing first order formulas over a ranked alphabet, and the basic functions and predicates over these types:

@@

\subsubsection{Example: a toy language}
In order to show a few simple examples of statements and proofs, we define a small language with very few symbols:
two constant symbols: $a$ and $b$, three function symbols $f$, $g$ and $h$  (of respective arity 1, 1 and 2), three propositional symbols $A$, $B$ and $C$, a one-place predicate symbol $P$, and a binary relational symbol $R$.

@@

\begin{remark}
  \label{rem:underscores}
  The constructors of types \texttt{Rel} and \texttt{Fun} are suffixed by an underscore, in order to reserve the names \texttt{a}, \texttt{f}, \texttt{h}, \texttt{R}, etc. to functions which build terms and formulas (please look at Sect~\ref{sec:fol-term-notations} and \ref{sec:fol-atomic-notations}).
\end{remark}

@@

\begin{remark}
This representation of terms uses mutually inductive data-types instead of list of terms. Please see also Remark~\ref{hydra:mutually-inductive-vs-lists}.
\end{remark}

\begin{remark}[Variables]
In O'Connor's formalization of first-order logic,  variables are 
just natural numbers, and the \emph{coercion} from
\texttt{nat} to \texttt{Term L} is the constructor (\texttt{@var $L$}).
Although other choices may be considered : PHOAS, de Bruijn indices, etc,  we still the data structures of~\cite{Goedel}, in order not to break long proof scripts which use this representation (please look at \cite{OConnor05} for a related discussion).

@@

\subsubsection{Examples}
\label{sect:folTermExamples}

Let us define some \gallina terms over our toy language, respecively corresponding to the terms $a$,
$f(a)$, and $h(f(a),a)$.

@@

\begin{remark}
The fourth ``term'' \texttt{t4} is not well formed, since the arity of $h$ is not respected\,\footnote{Strictly speaking, it's not a (well typed) term!}.
\end{remark}

@@

The type of first order formulas over $L$ is defined 
in~\href{../theories/html/hydras.Ackermann.fol.html}{Ackermann.fol} as an inductive data type.

@@

\begin{remark}
The type \texttt{Formula} (over $L$) has only five constructors,
respectively associated with term equalities $t_1=t_2$, atomic formulas of the form $(r\;t_1\;\dots\;t_n)$, implications $(F \arrow G)$, negations $({\sim}F)$, and universal quantifications $(\forall\,i, F)$.

@@

For instance, the the  term \texttt{(var 0)}
occurs freely and also inside the scope of a quantifier in the
formula \texttt{F4} \vpageref[above]{sect:folFormExamples}.

>>>>

This chapter is devoted to the presentation of  data structures for representing terms and first order formulas over a ranked alphabet, and the basic functions and predicates over these types, more precisely:

@@

\subsubsection{Example: $L$, a toy language}
In order to show a few simple examples of statements and proofs, we define a small language with very few symbols:
two constant symbols: $a$ and $b$, three function symbols $f$, $g$ and $h$  (of respective arity 1, 1 and 2), three propositional symbols $A$, $B$ and $C$, a one-place predicate symbol $P$, and a binary relational symbol $R$.

@@

\begin{remark}
  \label{rem:underscores}
  The constructors of types \texttt{Rel} and \texttt{Fun} are suffixed by an underscore, in order to reserve the names \texttt{a}, \texttt{f}, \texttt{h}, \texttt{R}, etc. to the functions which build terms and formulas (please look at Sect~\ref{sec:fol-term-notations} and \ref{sec:fol-atomic-notations}).
\end{remark}

@@

\begin{remark}
This representation of terms uses mutually inductive data-types instead of lists or vectors of terms. Please see also Remark~\ref{hydra:mutually-inductive-vs-lists}.
\end{remark}

\begin{remark}[Variables]
In O'Connor's formalization of first-order logic,  variables are 
just natural numbers, and the \emph{coercion} from
\texttt{nat} to \texttt{Term L} is the constructor (\texttt{@var $L$}).
Although other choices may be considered : PHOAS, de Bruijn indices, etc,  we still the data structures of~\cite{Goedel}, in order not to break long proof scripts which use this representation (please look at Section 2 of~\cite{OConnor05} for a related discussion).

@@

\subsubsection{Examples}
\label{sect:folTermExamples}
\label{sec:fol-term-notations}

Let us build a few \gallina terms over our toy language, 
respectively corresponding to the terms $a$,
$f(a)$, $h(f(a),a)$, and $h(f(v_0),g(v_1))$.

@@

The following ``term'' \texttt{t4} is not well formed, since the arity of $h$ is not respected\,\footnote{Strictly speaking, it's not a (well typed) term!}.

\inputsnippets{FolExamples/TermExamplesFail}

@@

The type of first order formulas over $L$ is defined 
in~\href{../theories/html/hydras.Ackermann.fol.html}{Ackermann.fol} as an inductive data type, with a limited set of basic constructions:
\emph{term equalities} $t_1=t_2$,
\emph{atomic propositions} $R\;t_1\;\dots\;t_n$, where $R$ is a relation symbol of arity $n$,
\emph{implications} $A \rightarrow B$,
\emph{negations} $\sim\,A$,
and \emph{universal quantifications} $\forall\,i,\;A$.

\noindent From~\href{../theories/html/hydras.Ackermann.fol.html}{Ackermann.fol}

@@

\begin{remark}

@@

For instance, the term \texttt{(var 0)}
occurs freely and also inside the scope of a quantifier in the
formula \texttt{F4} \vpageref[above]{sect:folFormExamples}.
